### 2022-10-05 paper/300-name-collisions.tex
Work on paper

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope.

@@

\section{Solutions}

@@

\subsection{Using scopes to distinguish names}

>>>>

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope. Note that the T in the type of the second T is ambiguous.

@@

\section{Alternative Solutions}

@@

Anytime that we introduce a new name in a type, we could check if the name already exists in the environment, and if it does, choose a different unique name. This approach is possible but tedious to implement in Statix. It requires a new relation to traverse through the type and rename. It also increases the time complexity, as constant traversals of the type are needed.

\section{Using scopes to distinguish names}

The solution we found to work best in the end is to change the definition of \verb|ID|. To be precise, at the grammar level we have two sorts, \verb|RID| is a "Raw ID", being just a string. \verb|ID| will have two constructors, one being \verb|Syn|, a syntactical \verb|ID|. The second one is \verb|ScopedName|, it is defined in statix, so there is no syntax for it. It will be generated by \verb|typeOfExpr|.

\begin{lstlisting}
context-free sorts ID
lexical sorts RID
context-free syntax
   ID.Syn = RID
signature constructors
  ScopedName : scope * RID -> ID
\end{lstlisting}

The \verb|ScopedName| constructor has a scope and a raw ID. The scope is used to uniquely identify the name. The main idea is that whenever we encounter a syntactical name, we replace it with a scoped name, so it unambiguous. The scope graph will never have a syntactical name in it. However, when querying the scope graph for a syntactical name, we return the lexically closest name.

\subsection{The example revisited}

In this section, we will step through the steps that happen during the type checking of the term above, with name collisions solved. To find the type, the following is evaluated, note that the names are now wrapped in a \verb|Syn| constructor:

\begin{lstlisting}
typeOfExpr(_, FnConstruct(Syn("T"), Type(), 
	FnConstruct(Syn("T"), Var(Syn("T")), Var(Syn("T")))))
\end{lstlisting}

\noindent
The name in the \verb|FnConstruct| is replaced with a scoped name. The scope of the name is the scope that the name is first defined in. We then type check the body with this scope.

\begin{lstlisting}
typeOfExpr(s1, FnConstruct(Syn("T"), Var(Syn("T")), Var(Syn("T"))))

(ScopedName(s1, T) : Type)
	       [s1]
\end{lstlisting}

\noindent
The same thing happens, the body of the \verb|FnConstruct| is typechecked with a new scope. Note that the type of the new \verb|T| now specifies which \verb|T| it means, so it is no longer ambiguous.

\begin{lstlisting}
typeOfExpr(s2, Var(Syn("T")))

(ScopedName(s1, T) : Type)     (ScopedName(s2, T) : Var(ScopedName(s1, T)))
	       [s1]<----------------------------------[s2]
\end{lstlisting}

Finally, we need to find the type of \verb|T|. This finds the lexically closest definition of \verb|T| (the one in s2), as defined earlier. The type of this \verb|T| is \verb|ScopedName(s1, T)|, which explicitly defined which \verb|T| it is. A name can now never shadow another name, since each scope uniquely identifies a name. The final type of the expression is now:

\begin{lstlisting}
FnType(ScopedName(s1, T), Type(), FnType(ScopedName(s2, T), Var(ScopedName(s1, T)), Var(ScopedName(s1, T))))
\end{lstlisting}

\subsection{Improving the readability of types}
